Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-03-31 article/article.Rmd
+ included chisq stats in article + revert to using glmm for analysis + set nmds max.stress to 0.25 to reduce dimensionality

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for a detailed explanation of data collection methods]. This data set was particularly well suited for our study because it contains genetic data for a diverse range of species that were collected using the same sampling regime. Additionally, the study area covers a broad geographic domain that encompasses  attitudinal gradients which provide substantial variation to examine the surrogates. As part of this project, data was collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus the data set contains information describing the genomic variation of individuals for each species present in every second grid cell. We used these grid cells as planning units to develop prioritizations. To reduce computational effort, we used only grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$). The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

@@

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of the species' intra-specific variation [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation, we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using surrogate-based targets ranging from `r (min(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % and `r (max(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target)*100)` % in increments of `r diff(rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.target[1:2])*100` % ($n = `r rapr.params.LST[[MODE]][['surrogacy.analysis']]$surrogate.optimal.replicates`$ per target). Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis.

To determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios, we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises. We used `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. We set surrogate- and genetic-based targets as `r (rapr.params.LST[[MODE]][['single.species']]$amount.target*100)` % to ensure that most of the respective variation was secured for each species. To accomodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]][['single.species']]$amount.replicates)` replicate prioritizations. Since all other contexts were likely to be associated with few optimal solutions, we generated only `r rapr.params.LST[[MODE]][['multi.species']]$genetic.replicates` replicates.

## Statistical analysis
We investigated the effectiveness of prioritizations generated using surrogate-based targets using generalized linear mixed-effects models [GLMMs; using a logit link and the $\texttt{lme4}$ R package; @r469]. To determine  if the environmental and geographic variables were effective surrogates, we fit a full model to the proportion of genetic variation that the first collection of prioritizations. This full model contained categorical predictor variables to indicate the surrogate-based targets used to generate them, the type of genetic variation (adaptive or neutral), and the interactions between these variables. To determine if surrogate-based targets improved the effectiveness of prioritizations under various scenarios, we fit a full model to the proportion of genetic variation that second collection of prioritizations secured. This second full model contained categorical predictor variables to indicate the scenario (single-species, multi-species, multi-species with cost), the combination of targets used to generate the prioritizations (amount-based targets, amount and surrogate-based targets, amount and genetic-based targets), the type of genetic variation measured (adaptive or neutral), and the interactions between them. All models contained intercept-level random effects to accommodate variation between species. We conducted a post-hoc analysis on each full model using Tukey contrasts with a Bonferroni correction [using the $\texttt{multcomp}$ R package; @r470].

>>>>

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set was particularly well suited for our study because it contains genetic data for a diverse range of species that were collected using the same sampling regime. Additionally, the study area covers a broad geographic domain that encompasses attitudinal gradients. As part of this project, data were collected for 27 alpine plant species using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species. Thus the data set contains information describing the genomic variation of individuals for each species present in every second grid cell. We used these grid cells as planning units to develop prioritizations. To reduce computational effort, we used only grid cells that contained at least a single individual ($n = `r nrow(grid.DF)`$). The majority of our analyses were performed using the \texttt{R} software package [version `r paste0(version$major, '.', version$minor)`; @r192].

@@

### Prioritizations
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of the species' intra-specific variation [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463]. Statistical analysis were conducted using the $\texttt{RVAideMemoire}$ R package [@rXXX].

To determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation (respectively), we generated a collection of single-species prioritizations. For each species, we generated two sets of prioritizations generated using increasing environmental and geographic targets (targets $\in$ `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target, collapse=', ')`). We calculated the proportion of adaptive genetic variation secured by the environmental-based prioritizations. We also calculated the proportion of neutral genetic variation secured by the geographic-based prioritizations. For each target, we counted the number of species that had a sufficient proportion of their genetic variation secured ($\geq$ `r rapr.params.LST[[MODE]]$surrogacy.analysis$genetic.target`). We used a $\chi^2$ test with Yates' continuity correction to examine whether the number of species with their genetic variation sufficiently preserved depended on the targets used to generate the prioritizations, and if the number of species sufficiently preserved given the same target differed depending on the type of genetic variation (adaptive vs. neutral) assessed. We then conducted post-hoc pairwise comparison analysis using $chi^2$ tests with Yates' continuity and Bonferroni corrections.

To determine if the addition of surrogate-based targets to conventional reserve selection methods can yield more effective prioritizations, we investigated the effectiveness of prioritizations that were generated using different types of targets. We generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets (environmental and geographic targets), and (3) both amount- and genetic-based targets (adaptive and neutral genetic variation targets). The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises. We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. We set the surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % since this was high enough to secure most of the genetic variation according to the surrogacy analysis.

We generated prioritizations using these three combinations of targets under different scenarios to determine if including surrogate-based targets is beneficial across a range of different conservation planning contexts. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. To determine if the benefits of surrogate-based targets are still evident when using a comprehensive set of species in a multi-species context, our second scenario involved generating multi-species prioritizations. To understand the effects of opportunity cost on the benefits of surrogate-based targets, our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost. Thus three types of prioritizations were generated for each scenario. Since multiple optimal solutions were available for the amount-only prioritizations in the single-species context, we generated `r rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates` prioritizations per species to accommodate variation in the proportion of genetic variation they secured.

After generating all the prioritizations, we then counted the number of species that had a sufficient proportion of their genetic variation secured ($\geq$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target`) in each of the prioritizations. We used a $\chi^2$ test with Yates' continuity correction to assess whether the number of species with a sufficient proportion of their genetic variation secured depended on the combination of targets used to generate the prioritization (amount-only, amount and surrogate-based targets, amount and genetic-based targets), the scenario under which the prioritization was generated (single species, multi-species, or multi-species with cost), and the type of genetic variation measured (adaptive or neutral). Finally, we conducted a post-hoc pairwise comparison analysis using $chi^2$ tests with Yates' continuity and Bonferroni corrections.

### 2016-03-31 article/figures.Rmd
+ included chisq stats in article + revert to using glmm for analysis + set nmds max.stress to 0.25 to reduce dimensionality

```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
correlation.PDF <- correlation.sub.DF %>%
	group_by(Surrogate.target, Type, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

correlation.PDF2 <- correlation.PDF %>%
	group_by(Type, Surrogate.target, Type) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>%
	mutate(Combination=as.character(paste0(Surrogate.target,'.',Type))) %>% 
	left_join(
		data.frame(
			letter=unname(correlation.fisher.cld$Letters),
			Combination=names(correlation.fisher.cld$Letters),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(value, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	left_join(
		data.frame(
			letter=unname(scenario.fisher.cld$Letters),
			Combination=names(scenario.fisher.cld$Letters)
		) %>% mutate(Combination=as.character(Combination)),
		by='Combination'
	) %>% data.frame()

>>>>

```{r, echo=FALSE, message=FALSE, fig.width=6, fig.height=4.5, fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. White boxplots show the proportion of genetic variation secured by prioritizations generated using surrogate-based targets. Grey boxplots show the proportion of genetic variation secured by randomly generated prioritizations with the same number of planning units as the surrogate-based prioritizations. Panel (a) shows the relationship between environmental surrogate targets used to generate a prioritization and the proportion of adaptive genetic variation it secures. Similarly, panel (b) shows that for geographic surrogate targets and neutral genetic variation. Box-plots show inter-specific variation in the proportion of genetic variation secured using different surrogate-based targets. Letters indicate significant differences ($P < 0.05$).'}
# prepare data
correlation.PDF <- correlation.sub.DF %>%
	group_by(Type, Surrogate.target, Method, Species) %>% 
	summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
	data.frame()

correlation.PDF2 <- correlation.PDF %>%
	group_by(Type, Surrogate.target, Method) %>% 
	summarize(genetic.held=max(genetic.held, na.rm=TRUE)+0.1) %>%
	ungroup() %>%
	mutate(Combination=as.character(paste0(Type,'.',Surrogate.target,'.',Method))) %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(correlation.GLHT)$mcletters$Letters),
			Combination=as.character(names(toupper(cld(correlation.GLHT)$mcletters$Letters))),
			stringsAsFactors=FALSE
		),
		by='Combination'
	) %>% data.frame()

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(value, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	left_join(
		data.frame(
			letter=unname(toupper(cld(scenario.GLHT)$mcletters$Letters)),
			Combination=as.character(names(toupper(cld(scenario.GLHT)$mcletters$Letters)))
		) %>% mutate(Combination=as.character(Combination)),
		by='Combination'
	) %>% data.frame()
